---
title: "R Notebook"
output:
html_notebook: default
---

<!-- Plot Figure 5: Capturing the emergence of resistance in lung cancer cohort.-->
<!-- Predicting the development of resistance to tyrosine kinase inhibitors (TKIs) in lung cancer patients. -->
```{r, setup, include=FALSE}
# Set working directory to the directory where "SCPO_submission" is cloned.
working_directory='/Users/sinhas8/SCPO_submission/'
knitr::opts_knit$set(root.dir = working_directory)
```

```{r}
onTarget=readRDS('Data/DepMapv12.RDS')
# Function to load a RMD file to another
ksource <- function(x, ...) {
  library(knitr)
  source(purl(x, output = tempfile()), ...)
}
ksource('Tools/step0A_functions_needed.Rmd')
ksource('Tools/step0B_functions_needed.Rmd')
genesUsed_toBuild=readRDS('Data/genesUsed_toBuild.RDS')
genes_across_scRNA_datasets_ofInterest=readRDS('Data/genes_across_scRNA_datasets_ofInterest.RDS')
```

<!-- Here, we preprocess and analyze Maynard et al dataset. -->
The data was provided by the authors as a box link which was downloaded to:
/data/Lab_ruppin/resources/single_cell_datasets/PRJNA591860/data_byAuthors/

We generated a symbolic link to the data in the working folder
ln -s /data/Lab_ruppin/resources/single_cell_datasets/PRJNA591860
/data/Lab_ruppin/vegesnar2/sc_drug_response/patient_validation/lung_cancer/PRJNA591860/data
<!-- patient data parsing has already been performed and objects are saved to save time. 
(Time Intensive and thus this is commented out) -->
```{r}
# data_id="PRJNA591860"
# expr="data/PRJNA591860_all_cells.RDS"
# met="data/PRJNA591860/data_byAuthors/NI05_all_epithelial_annotated_normal_and_tumor_Meta.txt"
# #Read files
# exp_matrix<-readRDS(expr)
# #epitheliali cells metadata only, exculdes the annotation of other cell types
# metadata<-read.table(met, sep="\t",stringsAsFactors=F,header=T, comment="")
# patient_sample_mappign=data.frame(metadata$sample_name,
#                                   metadata$patient_id)
# 
# celltype<-names(table(metadata$epi_anno_final, useNA = "ifany"))
# sample_names<-names(table(metadata$sample_name, useNA = "ifany"))
# patient_names<-names(table(metadata$patient_id, useNA = "ifany"))
# biopsy_names<-names(table(metadata$biopsy_site, useNA = "ifany"))
# stage_names<-names(table(metadata$primary_or_metastaic, useNA = "ifany"))
# 
# # subsettting malignant cells & metadata for malignant cells
# celltype_of_interest="tumor"
# malignant_metadata=metadata[which(metadata$epi_anno_final== celltype_of_interest),]
# malignant_cells<-malignant_metadata$cell_id
# 
# #Cells missing in expression matrix. Samples AZ_01 to AZ_05 are missing in the data. Alls samples starting with LT are present.
# missing=setdiff(malignant_cells,intersect(colnames(exp_matrix),malignant_cells))
# missing_metadata=malignant_metadata[match(missing,malignant_metadata$cell_id),]
# #Missing samples
# table(missing_metadata$sample_name)
# malignant_cells_parsed<-setdiff(malignant_cells,missing)
# 
# # Malignant cells expression.
# temp<-exp_matrix[,malignant_cells_parsed]
# PRJNA591860<- as.data.frame(as.matrix(temp))
# saveRDS(PRJNA591860,"data/PRJNA591860.RDS")
# PRJNA591860_sample_cell_names=lapply(sample_names, function(x){
#     samples_cells<-metadata$cell_id[which(metadata$sample_name==x)]
#  return(intersect(malignant_cells_parsed,samples_cells))})
# names(PRJNA591860_sample_cell_names)=sample_names
# saveRDS(PRJNA591860_sample_cell_names,"data/PRJNA591860_sample_cell_names.RDS")
```
<!-- Load Patient malignant cells expression and metadata -->
```{r}
PRJNA591860_sample_cell_names=readRDS("Data/PRJNA591860_sample_cell_names.RDS")
Total_cell_each_patient=sapply(PRJNA591860_sample_cell_names, length)
# Total patients with at least one malignant cell (N=39)
sum(Total_cell_each_patient>0)
PRJNA591860=readRDS("Data/PRJNA591860.RDS")
dim(PRJNA591860)
Maynard_Supp2_Demo=readxl::read_xlsx('Data/Maynard_Supp2_Demo.xlsx', sheet = 2)
```
<!-- Use genes used to build the model to normalize the scRNA-expression -->
```{r}
# Genes common between bulk, single cell and patient datasets which were used to build models
genesUsed_toBuild=Reduce(intersect,list(rownames(onTarget$expression_20Q4),
                                        rownames(onTarget$scRNA_complete),
                                        rownames(PRJNA591860)))
```
<!-- seurat_clustering given an expression matrix -->
```{r}
require(Seurat)
seurat_clustering <- function(input_matrix){
  lung_so <- CreateSeuratObject(counts = input_matrix,
                                project = "lungv0",
                                min.cells = 3,
                                min.features = 200)
  lung_so <- NormalizeData(lung_so, normalization.method = "LogNormalize", scale.factor = 10000)
  lung_so <- FindVariableFeatures(lung_so, selection.method = "vst", nfeatures = 2000)
  lung_so <- ScaleData(lung_so)
  lung_so <- RunPCA(lung_so, features = VariableFeatures(object = lung_so))
  lung_so <- FindNeighbors(lung_so, dims = 1:10)
  lung_so <- FindClusters(lung_so, resolution = 0.8)
  return(lung_so)
}
set.seed(1)
lung_so=seurat_clustering(PRJNA591860)
lung_so <- RunUMAP(lung_so, dims = 1:10)
EF11_panelA_umap_lung_dataset=DimPlot(lung_so, reduction = "umap")
ggsave(EF11_panelA_umap_lung_dataset, filename = 'Figures/Extended_Figure11.pdf', height=6, width=6)
ggsave(EF11_panelA_umap_lung_dataset, filename = 'Figures/Extended_Figure11.png', height=6, width=6)
# <!-- Create Clone ID for each cell-->
PRJNA591860_subClone_id=Idents(lung_so)
```
<!-- Load clone ID below in case you wanna skip seurat step -->
```{r}
# PRJNA591860_subClone_id=seurat_clustering(PRJNA591860)
# saveRDS(PRJNA591860_subClone_id, 'Data/subClone_id_PRJNA591860.RDS')
# PRJNA591860_subClone_id=readRDS('/Data/subClone_id_PRJNA591860.RDS')
```
<!-- Add Patient annotation -->
```{r}
PRJNA591860_subClone_id_df=data.frame(PRJNA591860_subClone_id)
patient_ann=PRJNA591860_sample_cell_names
# Patient and clone info for each cell
paitent_n_cloneAnn=lapply(1:length(PRJNA591860_sample_cell_names), function(x) {
  err_handle(data.frame(clone_id=PRJNA591860_subClone_id[match(PRJNA591860_sample_cell_names[[
    names(PRJNA591860_sample_cell_names)[x] ]], names(PRJNA591860_subClone_id))],
             patient_id=names(PRJNA591860_sample_cell_names)[x],
             cell_id= PRJNA591860_sample_cell_names[[ names(PRJNA591860_sample_cell_names)[x] ]]
  ))
  }
)
names(paitent_n_cloneAnn)=names(PRJNA591860_sample_cell_names)
paitent_n_cloneAnn_df=do.call(rbind, paitent_n_cloneAnn)
head(paitent_n_cloneAnn_df)
Cluster_abundance_in_patients=aggregate(cell_id~ clone_id+patient_id , paitent_n_cloneAnn_df, length)
Cluster_abundance_in_patients_list=split(Cluster_abundance_in_patients ,Cluster_abundance_in_patients$patient_id)
Cluster_abundance_in_patients_2plot=do.call(rbind, lapply(Cluster_abundance_in_patients_list, function(x) data.frame(x, proportion=x[,3]/sum(x[,3]))))
Cluster_abundance_in_patients_2plot$Treatment_Timepoint=Maynard_Supp2_Demo$`Treatement Timepoint`[match(Cluster_abundance_in_patients_2plot$patient_id, Maynard_Supp2_Demo$`Sample Name`)]
```
<!-- Subclonal Expression (approximation, sum over the cluster) -->
```{r}
paitent_n_cloneAnn=paitent_n_cloneAnn[names(unlist(sapply(paitent_n_cloneAnn,nrow)))]

PRJNA591860_subclone_meanExp=lapply(paitent_n_cloneAnn,function(x) 
  data.frame(sapply(split(x$cell_id, factor(x$clone_id)), function(y)
    rowMeans_if_one_row(data.frame(PRJNA591860[,y])) ))
         )
```
<!-- Obtain the drug response models -->
```{r}
file_list=list.files(path="Data/PRJNA591860/response_models/",full.names=TRUE)
model_file_list=file_list[grep("_model.RDS",file_list)]
DOI=unlist(lapply(model_file_list,function(x){unlist(strsplit(tail(unlist(strsplit(x,split="/")),1),split="_"))[1]}))
# Load the models for the drugs in the cohort
drugs_used=c("alectinib","ceritinib","crizotinib","dabrafenib","erlotinib","gemcitabine","lorlatinib","osimertinib","trametinib")
DOI_file_id=unlist(sapply(drugs_used, function(x) grep(x, model_file_list) ))
drug_models=lapply(DOI_file_id, function(x) readRDS(model_file_list[x]))
length(drug_models)
names(drug_models)=names(DOI_file_id)
```
<!-- Model Performance -->
```{r}
corr_matrix=cbind(
  bulk=unlist(lapply(drug_models,function(x)x$performance_in_bulk[2])),
  pseudo_bulk=unlist(lapply(drug_models,function(x)x$performance_in_pseudo_bulk[2])),
  sc=unlist(lapply(drug_models,function(x)x$performance_in_scRNA[2])) )
rownames(corr_matrix)=names(drug_models)

pval_matrix=cbind(
  bulk=unlist(lapply(drug_models,function(x)x$performance_in_bulk[1])),
  pseudo_bulk=unlist(lapply(drug_models,function(x)x$performance_in_pseudo_bulk[1])),
  sc=unlist(lapply(drug_models,function(x)x$performance_in_scRNA[1]))
)
rownames(pval_matrix)=names(drug_models)
#Number of cell lines with AUC values used in validation
TotalcellLines_used_in_validation=unlist(lapply(
  drug_models,function(x)length(na.omit(x$predVSgroundTruth$pred_gt_mscRNA$Observed))))
```
<!-- Select the significant drug models -->
```{r}
significant_drugs_names=names(which(corr_matrix[,3]>0.3 & pval_matrix[,3]<0.05)) #
#Drug models from PERCEPTION
significant_drugs_models=drug_models[significant_drugs_names]
#In control mode:: Repeat with (Bulk_models_list) from "Bulk_based_only_model.Rmd"
# significant_drugs_names=Bulk_models_list
# significant_drugs_names=Bulk_models_list_rf
# significant_drugs_models=Bulk_models_list
```
<!-- #COmpute Improvement Score-->
<!-- Step 1: Monotherapy Killing for each clone-->
```{r}
PRJNA591860_subclone_meanExp_rnorm=lapply(PRJNA591860_subclone_meanExp, function(x) 
  err_handle(data.frame(rank_normalization_mat(data.frame(x[genesUsed_toBuild,])), 
                        row.names = genesUsed_toBuild) ) )
killing_eachClone_list= lapply(PRJNA591860_subclone_meanExp_rnorm, function(x)
  err_handle(killing_in_each_dataset(infunc_scRNAseq_dataset_rnorm=x,
                                     infunc_GOI=genesUsed_toBuild,
                                     infun_model_list=significant_drugs_models)) )

killing_eachClone_list_wd_paitentID=lapply(names(killing_eachClone_list), function(x) cbind(x,
                                                        killing_eachClone_list[[x]]))
names(killing_eachClone_list_wd_paitentID)=names(killing_eachClone_list)
Killing_at_clone_level=do.call(rbind, lapply(names(Cluster_abundance_in_patients_list), function(x)
  data.frame(CloneID = Cluster_abundance_in_patients_list[[x]]$clone_id,
             killing_eachClone_list_wd_paitentID[[x]]) ))
```
<!-- Plot EF11_panelB -->
```{r}
EF11_panelB <- ggplot(Cluster_abundance_in_patients_2plot,
                 aes(fill=clone_id, y=proportion, x=patient_id)) + 
    geom_bar(position="stack", stat="identity")+
  theme_bw(base_size = 15)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = 'top')+
  facet_grid(.~Treatment_Timepoint, shrink = T, scales = "free", space = "free_x")+
  labs(y='Clone\nProportion')
ggsave(EF11_panelB, filename = 'Figures/EF11_panelB.pdf', height=4, width=6)
ggsave(EF11_panelB, filename = 'Figures/EF11_panelB.png', height=4, width=6)
```
<!-- Plot Killing at clone level for Erlotinib -->
```{r}
Killing_at_clone_level$Treatment=Maynard_Supp2_Demo$Treatment[
  match(Killing_at_clone_level$x, Maynard_Supp2_Demo$`Sample Name`)]
Killing_at_clone_level$Response=Maynard_Supp2_Demo$`Treatement Timepoint`[
  match(Killing_at_clone_level$x, Maynard_Supp2_Demo$`Sample Name`)]
erlotinib_id=grep('erlotinib',Killing_at_clone_level$Treatment)
osimertinib_id=grep('osimertinib',Killing_at_clone_level$Treatment)
dabrafenib_id=grep('dabrafenib',Killing_at_clone_level$Treatment)
trametinib_id=grep('trametinib',Killing_at_clone_level$Treatment)
Killing_at_clone_level$erlotinib=as.numeric(Killing_at_clone_level$erlotinib)
Killing_at_clone_level$osimertinib=as.numeric(Killing_at_clone_level$osimertinib)
Killing_at_clone_level$gemcitabine=as.numeric(Killing_at_clone_level$gemcitabine)
Killing_at_clone_level$dabrafenib=as.numeric(Killing_at_clone_level$dabrafenib)
Killing_at_clone_level$trametinib=as.numeric(Killing_at_clone_level$trametinib)

EF11_panel_C=ggplot(Killing_at_clone_level[erlotinib_id,], aes(y=erlotinib, x=CloneID))+
  geom_point(aes(color= erlotinib, size=erlotinib))+
  geom_segment( aes(x=CloneID,
                    xend=CloneID, y=0, 
                    yend=erlotinib), color="black")+
  facet_grid(.~Response+x,
             scales = 'free_x', shrink = T,
             drop = T,  space = "free_x", switch = 'x',
             as.table = T)+
  theme_bw(base_size = 11)+
  scale_colour_viridis_c()+
  theme(legend.position = 'top',
        strip.placement = 'outside',
        strip.background = element_rect(fill="white", size=1, color="white"),
        strip.text = element_text(angle=0,
                                  color=c(rep(hue_pal()(2)[1],7),rep(hue_pal()(2)[2],21))
  )
        )+
  labs(x='Clones', y='Predicted Viability of\nErlotinib (z-score)',
       color= 'Predicted Viability', size='Predicted Viability')

ggsave(EF11_panel_C, filename = 'Figures/EF11_panel_C.pdf', height=5, width=9)
ggsave(EF11_panel_C, filename = 'Figures/EF11_panel_C.png', height=5, width=9)
```
<!-- Plot Killing at clone level for osimertinib -->
```{r}
EF11_panel_D=ggplot(Killing_at_clone_level[osimertinib_id,], aes(y=osimertinib, x=CloneID))+
  geom_point(aes(color= osimertinib, size=osimertinib))+
  geom_segment( aes(x=CloneID,
                    xend=CloneID, y=0, 
                    yend=osimertinib), color="black")+
  facet_grid(.~Response+x,
             scales = 'free_x', shrink = T,
             drop = T,  space = "free_x", switch = 'x',
             as.table = T)+
  theme_bw(base_size = 11)+
  scale_colour_viridis_c()+
  theme(legend.position = 'top',
        strip.placement = 'outside',
        strip.background = element_rect(fill="white", size=1, color="white"),
        strip.text = element_text(angle=0,
                                  color=c(rep(hue_pal()(2)[1],7),rep(hue_pal()(2)[2],21))
  )
        )+
  labs(x='Clones', y='Predicted Viability of\nosimertinib (z-score)',
       color= 'Predicted Viability', size='Predicted Viability')
```
<!-- Plot Killing at clone level for dabrafenib -->
```{r}
EF11_panel_F=ggplot(Killing_at_clone_level[dabrafenib_id,], aes(y=dabrafenib, x=CloneID))+
  geom_point(aes(color= dabrafenib, size=dabrafenib))+
  geom_segment( aes(x=CloneID,
                    xend=CloneID, y=0, 
                    yend=dabrafenib), color="black")+
  facet_grid(.~Response+x,
             scales = 'free_x', shrink = T,
             drop = T,  space = "free_x", switch = 'x',
             as.table = T)+
  theme_bw(base_size = 11)+
  scale_colour_viridis_c()+
  theme(legend.position = 'top',
        strip.placement = 'outside',
        strip.background = element_rect(fill="white", size=1, color="white"),
        strip.text = element_text(angle=0,
                                  color=c(rep(hue_pal()(2)[1],7),rep(hue_pal()(2)[2],21))
  )
        )+
  labs(x='Clones', y='Predicted Viability of\ndabrafenib (z-score)',
       color= 'Predicted Viability', size='Predicted Viability')
```
<!-- Plot Killing at clone level for trametinib -->
```{r}
EF11_panel_G=ggplot(Killing_at_clone_level[trametinib_id,], aes(y=trametinib, x=CloneID))+
  geom_point(aes(color= trametinib, size=trametinib))+
  geom_segment( aes(x=CloneID,
                    xend=CloneID, y=0, 
                    yend=trametinib), color="black")+
  facet_grid(.~Response+x,
             scales = 'free_x', shrink = T,
             drop = T,  space = "free_x", switch = 'x',
             as.table = T)+
  theme_bw(base_size = 11)+
  scale_colour_viridis_c()+
  theme(legend.position = 'top',
        strip.placement = 'outside',
        strip.background = element_rect(fill="white", size=1, color="white"),
        strip.text = element_text(angle=0,
                                  color=c(rep(hue_pal()(2)[1],7),rep(hue_pal()(2)[2],21))
  )
        )+
  labs(x='Clones', y='Predicted Viability of\ntrametinib (z-score)',
       color= 'Predicted Viability', size='Predicted Viability')
```
<!-- Plot the Whole EF 11 -->
```{r}
EF11 <- grid.arrange(EF11_panelA_umap_lung_dataset,
                     EF11_panelB,
                     EF11_panel_C,
                     EF11_panel_F,
                     EF11_panel_D,
                     EF11_panel_G,
                     layout_matrix = rbind(c(1, 2, 2),c(1, 2, 2),c(3), c(4),c(5), c(6)))
ggsave(EF11, filename = 'Figures/Extended_Figure11.pdf', height=16, width=12)
ggsave(EF11, filename = 'Figures/Extended_Figure11.png', height=16, width=12)
```
<!-- Step 2: Patient level killing (colMax) -->
```{r}
# Convert list to df
killing_eachClone_list_wd_paitentID_df=do.call(rbind, killing_eachClone_list_wd_paitentID)
# Scale the killing
killing_eachClone_scaled=data.frame(Patient_ID=killing_eachClone_list_wd_paitentID_df[,1],
           apply(killing_eachClone_list_wd_paitentID_df[,-1], 2, function(x) scale(as.numeric(x))))
# Convert back to list
killing_eachClone_scaled_list=split(killing_eachClone_scaled[,-1],
                                    killing_eachClone_scaled$Patient_ID)
```
<!-- Compute Patient response from Clone response -->
```{r}
# Rule: Most resistant clone best reflects the clinical response
Killing_patient_level=sapply(killing_eachClone_scaled_list, colMax)

# Rule: Most resistant clone best reflects the clinical response
# Killing_patient_level=sapply(killing_eachClone_scaled_list, colMeans)
Killing_patient_level=data.frame(Killing_patient_level)
```
<!-- Checkpoint -->
```{r}
# saveRDS(Killing_patient_level, 'Data/PRJNA591860_patient_level_killing.RDS')
Killing_patient_level=readRDS('Data/PRJNA591860_patient_level_killing.RDS')
```
<!-- Load the demographics provided -->
```{r}
Killing_patient_level=t(Killing_patient_level)
Maynard_Supp2_Demo=readxl::read_xlsx('/Users/sinhas8/Project_scRNAbased_drugCombination/Data/Maynard_Supp2_Demo.xlsx', sheet = 2)
sort(table(Maynard_Supp2_Demo$`Patient ID`))
Killing_patient_level_ann_biopsy_timepoint=cbind(Killing_patient_level, 
Maynard_Supp2_Demo[match(rownames(Killing_patient_level),
                         Maynard_Supp2_Demo$`Sample Name`),])
```
<!-- Hypothesis: Whether the timepoint of tissue collection is positively correlated vs pred viab -->
<!-- Step 1: Compute each drug Treatnaive killing -->
```{r}
treatment_naive_samples=Killing_patient_level_ann_biopsy_timepoint$`Treatement Timepoint`=='TN'
significant_drugs_names=colnames(Killing_patient_level)
treatment_naive_Killing_each_drug=Killing_patient_level_ann_biopsy_timepoint[treatment_naive_samples,significant_drugs_names]
treatment_naive_Killing_each_drug_mean=colMeans(treatment_naive_Killing_each_drug)
treatment_naive_Killing_each_drug_mean
```
<!-- add the needed basline to the dataframe-->
```{r}
table(Killing_patient_level_ann_biopsy_timepoint$Treatment)
sort(unique(Killing_patient_level_ann_biopsy_timepoint$`Treatement Timepoint`))
progressive_diseases_samples=Killing_patient_level_ann_biopsy_timepoint$`Treatement Timepoint`!='TN'
Killing_patient_level_ann_biopsy_timepoint_Post=Killing_patient_level_ann_biopsy_timepoint[progressive_diseases_samples,]
Killing_patient_level_ann_biopsy_timepoint_Post$Treatment[grep('erlotinib',Killing_patient_level_ann_biopsy_timepoint_Post$Treatment)]='erlotinib'
Killing_patient_level_ann_biopsy_timepoint_Post$Treatment[grep('dabrafenib',Killing_patient_level_ann_biopsy_timepoint_Post$Treatment)]='dabrafenib'
Killing_patient_level_ann_biopsy_timepoint_Post$Treatment[grep('trametinib',Killing_patient_level_ann_biopsy_timepoint_Post$Treatment)]='trametinib'
Killing_patient_level_ann_biopsy_timepoint_Post=Killing_patient_level_ann_biopsy_timepoint_Post[
  Killing_patient_level_ann_biopsy_timepoint_Post$Treatment %in%
    names(treatment_naive_Killing_each_drug_mean),]
# add the needed basline
Killing_patient_level_ann_biopsy_timepoint_Post$respective_baseline=
  treatment_naive_Killing_each_drug_mean[
  match(Killing_patient_level_ann_biopsy_timepoint_Post$Treatment,
        names(treatment_naive_Killing_each_drug_mean))]
```
<!-- Extract each sample differential change and days of treatment -->
```{r}
DOI=unique(Killing_patient_level_ann_biopsy_timepoint_Post$Treatment)
dim(Killing_patient_level_ann_biopsy_timepoint_Post)

dim(Killing_patient_level_ann_biopsy_timepoint_Post)
Killing_patient_level_ann_biopsy_timepoint_Post
x=DOI[1]
Viab_preANDpost_treatment=lapply(DOI, function(x){
  infunc_df=Killing_patient_level_ann_biopsy_timepoint_Post[
    Killing_patient_level_ann_biopsy_timepoint_Post$Treatment %in% x,]
  
  
  cbind(resp=infunc_df[,x], 
        infunc_df[,c('Sample Name','Patient ID', colnames(infunc_df)[16:24])],
        Change_resp=infunc_df[,x] - infunc_df[,24])
  } )

Viab_preANDpost_treatment_integrated=do.call(rbind, Viab_preANDpost_treatment)
```
<!-- Correlation btw resistance and days of treatment -->
```{r}
resistant_id=Viab_preANDpost_treatment_integrated$`Treatement Timepoint`=='PD'
cor.test(
  as.numeric(Viab_preANDpost_treatment_integrated$`Treatment start to tissue acquisition (days)`)[resistant_id],
  as.numeric(Viab_preANDpost_treatment_integrated$Change_resp)[resistant_id],
  method = 's')
```
<!--Plot Correlation btw resistance and days of treatment for PD and RD-->
```{r}
Viab_preANDpost_treatment_integrated$biopsy_from_treatment=as.numeric(Viab_preANDpost_treatment_integrated$`Treatment start to tissue acquisition (days)`)
Viab_preANDpost_treatment_integrated$`Treatement Timepoint`=factor(Viab_preANDpost_treatment_integrated$`Treatement Timepoint`)
dim(Viab_preANDpost_treatment_integrated)

fig5A <- ggplot(Viab_preANDpost_treatment_integrated,
       aes(x=Change_resp,
           y=biopsy_from_treatment,
           fill=`Treatement Timepoint`,
           color=`Treatement Timepoint`))+
  geom_point()+
  stat_smooth(method = 'lm')+
  theme_bw(base_size = 20)+
  labs(y='Treatment Length', x="Extent of resistance\n from the baseline")
ggsave(fig5A,
       filename = 'Figures/Figure5A.pdf', height = 4, width = 8)
ggsave(fig5A,
       filename = 'Figures/Figure5A.png',height = 4, width = 8)
colnames(Viab_preANDpost_treatment_integrated)[1]='PERCEPTION Extent of Response'
write.csv(Viab_preANDpost_treatment_integrated,'Data/TableS5.csv')
```
<!-- Stratification -->
<!-- Figure 5B -->
```{r}
rocobj <- roc(response = Viab_preANDpost_treatment_integrated$`Treatement Timepoint`,
              predictor = Viab_preANDpost_treatment_integrated$Change_resp)
panelB <- ggroc(smooth(rocobj))+
  theme_bw(base_size = 15) +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed")+
  annotate("text", x=0.2, y=0.1, label= paste('AUC=',round(rocobj$auc,3)), size=5)
ggsave(panelB,
       filename = 'Figures/ROC_PDvsR.png',
       height = 4, width = 4)
ggsave(panelB,
       filename = 'Figures/ROC_PDvsR.pdf',
       height = 4, width = 4)
```
<!-- Plot Figure 5D -->
<!-- Hypothesis: Whether clones that are more resistant more likely to increase vs the clones that are senstitive -->
<!-- Load predicted killing at a clone level -->
```{r}
killing_n_abundance_clone_keyDrugs=readRDS('Data/lung_killing_n_abundance_clone_keyDrugs.RDS')
```
<!-- Preprocessing needed -->
```{r}
clones_with_more_than_1points=killing_n_abundance_clone_keyDrugs$CloneID %in% 
  names(which(table(killing_n_abundance_clone_keyDrugs$CloneID)>1))
killing_n_abundance_clone_keyDrugs=killing_n_abundance_clone_keyDrugs[clones_with_more_than_1points,]
killing_n_abundance_clone$treatment_time=as.numeric(killing_n_abundance_clone$treatment_time)
mean_killing_each_clone=data.frame(aggregate(killing_n_abundance_clone_keyDrugs$respective_response ~
                                               killing_n_abundance_clone_keyDrugs$CloneID, FUN = mean))
colnames(mean_killing_each_clone)=c('cloneID', 'mean_killing')
mean_killing_each_clone$mean_killingQ=xtile(mean_killing_each_clone$mean_killing, 2)
killing_n_abundance_clone_keyDrugs$Whether_resistant=mean_killing_each_clone$mean_killingQ[match( killing_n_abundance_clone_keyDrugs$CloneID, mean_killing_each_clone$cloneID)]
```
<!-- Linear regression to identify the slope of abundance vs time plot -->
```{r}
all_available_clone=unique(killing_n_abundance_clone_keyDrugs$CloneID)
# Linear Regression of Proportion vs treatment time
lm_propVStime=sapply(all_available_clone, function(x) {
  clone_of_interest=killing_n_abundance_clone_keyDrugs$CloneID==x
  killing_n_abundance_clone_keyDrugs_subset=killing_n_abundance_clone_keyDrugs[clone_of_interest,]
  summary(lm(killing_n_abundance_clone_keyDrugs_subset$proportion~
       killing_n_abundance_clone_keyDrugs_subset$treatment_time))
  }
  )

# Get coefficient & P
lm_propVStime_trimmed=sapply(lm_propVStime, function(x) x$coefficients[2,c(1,4)])
colnames(lm_propVStime_trimmed)=as.character(all_available_clone)
lm_propVStime_trimmed=as.data.frame(t(lm_propVStime_trimmed))
lm_propVStime_trimmed$mean_killing=mean_killing_each_clone$mean_killing[
  match(rownames(lm_propVStime_trimmed), mean_killing_each_clone$cloneID)]
lm_propVStime_trimmed$mean_killingQ=mean_killing_each_clone$mean_killingQ[
  match(rownames(lm_propVStime_trimmed), mean_killing_each_clone$cloneID)]
lm_propVStime_trimmed$mean_killingQ=factor(lm_propVStime_trimmed$mean_killingQ, labels= c('Bottom 50%\n resistant', 'Top 50% \nresistant'))

figure5D <- ggplot(lm_propVStime_trimmed, aes(y=Estimate, x=factor(mean_killingQ)))+
  geom_boxplot()+
  stat_compare_means(comparisons = combn(unique(lm_propVStime_trimmed$mean_killingQ), 2, simplify = F),method.args = list(alternative='l'))+
  theme_bw(base_size = 13)+
  labs(x='Clones', y='Rate of Emergence with\ntime elapsed from the\nstart of the treatment')
ggsave(filename = 'Figures/figure5Dv0.pdf',figure5D,  width = 3, height = 3)
```
<!-- Plot Extended_Figure_8 -->
<!-- Correlation btw resistance and days of treatment within each treatment-->
```{r}
only_PD=Viab_preANDpost_treatment_integrated$`Treatement Timepoint`=='PD'
Extended_Figure_8 <-  ggplot(Viab_preANDpost_treatment_integrated[only_PD,],
       aes(x=Change_resp,
           y=biopsy_from_treatment,
           fill=Treatment,
           color=Treatment))+
  geom_point()+
  stat_smooth(method = 'lm')+
  theme_bw(base_size = 20)+
  labs(y='Biopsies Days from \nthe start of treatment',
       x="Extent of Resistance")
ggsave(Extended_Figure_8,
       filename = 'Figures/Extended_Figure_8.pdf',
       height = 5, width = 7)
```
<!-- Correlation btw resistance and days of treatment within each prior treatment category-->
```{r}
Viab_preANDpost_treatment_integrated$`Treatment Hx`[grep('on firs',Viab_preANDpost_treatment_integrated$`Treatment Hx`)]='on firstline TKI'

Extended_Figure_8B <- ggplot(Viab_preANDpost_treatment_integrated,
       aes(x=Change_resp,
           y=biopsy_from_treatment,
           fill=`Treatment Hx`,
           color=`Treatment Hx`))+
  geom_point()+
  stat_smooth(method = 'lm')+
  theme_bw(base_size = 20)+
  labs(y='Biopsies Days from \nthe start of treatment',
       x="∆Viability differential\n from baseline")
ggsave(Extended_Figure_8B,
       filename = 'Figures/Extended_Figure_8B.png',
       height = 5, width = 7)
ggsave(Extended_Figure_8B,
       filename = 'Figures/Extended_Figure_8B.pdf',
       height = 5, width = 7)

```
<!-- Correlation btw resistance and days of treatment within each patient-->
```{r}
table(Viab_preANDpost_treatment_integrated$`Patient ID`[only_PD])
Viab_preANDpost_treatment_integrated[grep('TH179', Viab_preANDpost_treatment_integrated$`Patient ID`),]
Viab_preANDpost_treatment_integrated$`Patient ID`=gsub('_TAT','',Viab_preANDpost_treatment_integrated$`Patient ID`)
sort(table(Viab_preANDpost_treatment_integrated$`Patient ID`))
POI=names(which(sort(table(Viab_preANDpost_treatment_integrated$`Patient ID`[only_PD]))>1))
POI_id= Viab_preANDpost_treatment_integrated$`Patient ID` %in% POI
Extended_Figure_8C<- ggplot(Viab_preANDpost_treatment_integrated[POI_id,],
       aes(x=Change_resp,
           y=biopsy_from_treatment,
           fill=`Patient ID`,
           color=`Patient ID`))+
  geom_point()+
  stat_smooth(method = 'lm')+
  theme_bw(base_size = 20)+
  labs(y='Biopsies Days from \nthe start of treatment',
       x="∆Viability differential\n from baseline")
ggsave(Extended_Figure_8C,
       filename = 'Figures/Extended_Figure_8C.png',
       height = 5, width = 7)
ggsave(Extended_Figure_8C,
       filename = 'Figures/Extended_Figure_8C.pdf',
       height = 5, width = 7)

# Leveraging all the data to ask whether within patient - resistance increases with time.
summary(lm(Change_resp ~ 
             biopsy_from_treatment:factor(`Treatement Timepoint`),
   data = Viab_preANDpost_treatment_integrated))
```

```{r}
remove_NA_id=Viab_preANDpost_treatment_integrated$`Primary/Metastatic`!='NA'
Extended_Figure_8D<- ggplot(na.omit(Viab_preANDpost_treatment_integrated[only_PD & remove_NA_id,]),
       aes(x=Change_resp,
           y=biopsy_from_treatment,
           fill=`Primary/Metastatic`,
           color=`Primary/Metastatic`))+
  geom_point()+
  stat_smooth(method = 'lm')+
  theme_bw(base_size = 20)+
  labs(y='Biopsies Days from \nthe start of treatment',
       x="∆Viability differential\n from baseline")+
  scale_fill_manual(values=c("#999999", "#E69F00"))
ggsave(Extended_Figure_8D,
       filename = 'Figures/Extended_Figure_8D.png',
       height = 5, width = 7)
ggsave(Extended_Figure_8D,
       filename = 'Figures/Extended_Figure_8D.pdf',
       height = 5, width = 7)
table(Viab_preANDpost_treatment_integrated$`Biopsy Type`)
```
<!-- Panel 5F, main text -->
# Analyzing Cross resistancegith
```{r}
sort(table(Killing_patient_level_ann_biopsy_timepoint_Post$Treatment))
Cross_resistance_mat=Killing_patient_level_ann_biopsy_timepoint_Post
consider_PD_only=Cross_resistance_mat$`Treatement Timepoint`=='PD'
Cross_resistance_mat=Cross_resistance_mat[consider_PD_only,]
Killing_patient_level_ann_biopsy_timepoint_Post$Treatment=='dabrafenib'
treated_wd_erlotinib=Cross_resistance_mat$Treatment=='erlotinib'
treated_wd_osimertinib=Cross_resistance_mat$Treatment=='osimertinib'
treated_wd_dabrafenib=Cross_resistance_mat$Treatment=='dabrafenib'
treated_wd_trametinib=Cross_resistance_mat$Treatment=='trametinib'
table(treated_wd_trametinib)
Cross_resistance_mat$biopsy_from_treatment=
  as.numeric(Cross_resistance_mat$`Treatment start to tissue acquisition (days)`)
```
<!-- Cross-resistance strength and significnace. -->
```{r}
CR_correlation_strength=data.frame(
  Erlotinib_Treatment_Resistance=sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_erlotinib,]$erlotinib,
         Cross_resistance_mat[treated_wd_erlotinib,x])[c(4)])),
dabrafenib_Treatment_Resistance=sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_dabrafenib,]$dabrafenib,
         Cross_resistance_mat[treated_wd_dabrafenib,x])[c(4)])),
trametinib_Treatment_Resistance=err_handle(sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_trametinib,]$trametinib,
         Cross_resistance_mat[treated_wd_trametinib,x])[c(4)])))
)
CR_correlation_significance=data.frame(
  Erlotinib_Treatment_Resistance=sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_erlotinib,]$erlotinib,
         Cross_resistance_mat[treated_wd_erlotinib,x])[c(3)])),
dabrafenib_Treatment_Resistance=sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_dabrafenib,]$dabrafenib,
         Cross_resistance_mat[treated_wd_dabrafenib,x])[c(3)])),
trametinib_Treatment_Resistance=sapply(significant_drugs_names, function(x)
       unlist(cor.test(Cross_resistance_mat[treated_wd_trametinib,]$trametinib,
         Cross_resistance_mat[treated_wd_trametinib,x])[c(3)]))
)

```
<!-- Panel 5F: Plot the CR_corr_matrix -->
```{r}
CR_correlation_strength=as.matrix(CR_correlation_strength)
colnames(CR_correlation_strength)=c('Erlotinib',
                                    'Dabrafenib',
                                    'Trametinib')
rownames(CR_correlation_strength)=c('Dabrafenib',
                                    'Erlotinib',
                                    'Gemcitabine',
                                    'Osimertinib',
                                    'Trametinib')
CR_correlation_significance=as.matrix(CR_correlation_significance)
rownames(CR_correlation_significance)=rownames(CR_correlation_strength)
colnames(CR_correlation_significance)=colnames(CR_correlation_strength)

pdf(file = 'Figures/cross_resistance_correlation.pdf', width = 4.5, height=4)
corrplot(CR_correlation_strength,
         p.mat = CR_correlation_significance,
         addCoef.col ='black',
         insig = 'blank',
         sig.level = 0.2,
         number.cex = 1,
         tl.cex = 0.9,
         cl.cex = 0.75, cl.align.text = 'l')
dev.off()
```
<!-- Panel 5F -->
```{r}
P1 <- ggplot(Cross_resistance_mat[treated_wd_erlotinib,],
       aes(y=erlotinib, x=gemcitabine))+
  geom_point()+
  theme_bw(base_size = 20)+
  stat_smooth(method='lm')+
  labs(y='Erlotinib',
       x='gemcitabine')+
  stat_cor(size=5)
P2 <- ggplot(Cross_resistance_mat[treated_wd_erlotinib,],
       aes(y=erlotinib, x=dabrafenib))+
  geom_point()+
  theme_bw(base_size = 20)+
  stat_smooth(method='lm')+
  labs(y='Erlotinib',
       x='dabrafenib')+
  stat_cor(size=5)
P3 <- ggplot(Cross_resistance_mat[treated_wd_erlotinib,],
       aes(y=erlotinib, x=trametinib))+
  geom_point()+
  theme_bw(base_size = 20)+
  stat_smooth(method='lm')+
  labs(y='Erlotinib',
       x='Trametinib')+
  stat_cor(size=5)
P4 <- ggplot(Cross_resistance_mat[treated_wd_dabrafenib,],
       aes(y=dabrafenib, x=gemcitabine))+
  geom_point()+
  theme_bw(base_size = 20)+
  stat_smooth(method='lm')+
  labs(x='Gemcitabine',
       y='Dabrafenib')+
  stat_cor(size=5)
ggsave(ggarrange(P1, P2, P3, P4),
       filename = 'Figures/Figure_Cross_Resistance.pdf', height=6, width=6)
```

```{r}
dabrafenib=rbind(data.frame(Pred_Viability=treatment_naive_Killing_each_drug_mean['dabrafenib'],
                 Day=0, Point='pre-treatment'),
      data.frame(
        Pred_Viability=Cross_resistance_mat$dabrafenib[
          Cross_resistance_mat$Treatment=='dabrafenib'],
        Day=as.numeric(
          Cross_resistance_mat$`Treatment start to tissue acquisition (days)`[Cross_resistance_mat$Treatment=='dabrafenib']), 
         Point='Post-treatment'
      ) )
```

```{r}
dabrafenib_expanded=rbind(data.frame(Pred_Viability=treatment_naive_Killing_each_drug$dabrafenib,
                 Day=0, Point='pre-treatment'),
      data.frame(
        Pred_Viability=Cross_resistance_mat$dabrafenib[
          Cross_resistance_mat$Treatment=='dabrafenib'],
        Day=as.numeric(
          Cross_resistance_mat$`Treatment start to tissue acquisition (days)`[Cross_resistance_mat$Treatment=='dabrafenib']), 
         Point='Post-treatment'
      ) )

```
<!-- Panel 5D -->
```{r}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

df2 <- data_summary(dabrafenib_expanded, varname="Pred_Viability", 
                    groupnames=c("Day"))
df2$Sample_type='Post-Treatment'
df2$Sample_type[df2$Day==0]='Pre-Treatment'
ggplot(df2, aes(x=Day, y=Pred_Viability)) + 
  geom_line()+
  geom_pointrange(aes(ymin=Pred_Viability-sd, ymax=Pred_Viability+sd, color=Sample_type))


ggplot(df2, aes(x=Day, y=Pred_Viability)) + 
  geom_line()+
  geom_pointrange(aes(ymin=Pred_Viability-sd, ymax=Pred_Viability+sd, color=Sample_type))

df2$sd[is.na(df2$sd)]=0
panel_C <- ggplot(df2, aes(x=Day, y=Pred_Viability)) + 
  geom_line()+
  geom_point()+
  geom_errorbar(aes(ymin=Pred_Viability-sd, ymax=Pred_Viability+sd), width=20)+
  theme_bw(base_size = 15)+
  labs(x='Elapsed Treatment Time', y='Predicted Viability (z)')
ggsave(panel_C,
       filename = 
         'Figures/Figure7C_onePatient.pdf',
       height = 4, width = 4)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

