---
title: "R Notebook"
output:
html_notebook: default
---

<!-- Below functions are needed to build, train, and test PERCEPTION model in cell lines & patients -->
# This could be divided into two steps: 
# 1. Initial training in cell lines using bulk-expression and drug response.
# 2. Hyperparameter tuning using sc-expression in cell lines
<!-- Step 1A: Identify genes with bulk-expression are significantly correlated with drugs response -->
```{r}
# This function takes, drugname, cancer type, all gene features, response measure (AUC vs viability), which cancer type cell lines to exclude from analysis.
feature_ranking_bulk<-function(infunc_drugName=DOI[1],
                               infunc_cancerType='PanCan',
                               infunc_GOI=genesUsed_toBuild,
                               resp_measure_mode='AUC',
                               exclude_cancer='lung'){
  # Choose the resp matrix
  if(resp_measure_mode=='AUC'){
    # Mode 1: All HTS
    # infun_response_matrix=onTarget$secondary_prism
    # infunc_response=infun_response_matrix[
    #   onTarget$drugCategory$broad_id_trimmed[match(infunc_drugName, onTarget$drugCategory$name)],]
    # Mode 2: Priortize MTS
    infun_response_matrix=onTarget$secondary_prism
    infunc_drugName_id=which(stripall2match(onTarget$secondary_screen_drugAnnotation$CommonName)==infunc_drugName)
    infunc_response=onTarget$secondary_prism[infunc_drugName_id,]
    
    # is.matrix(infunc_response) - denotes if there exists a matrix with multiple screen
    # If there are multiple measures of response;
    if(is.matrix(infunc_response)){
      # Choose screen of interest
      # Rule: Priority, MTS over HTS; MTS with least number of NAs
      infunc_drugName_screenID=onTarget$secondary_screen_drugAnnotation$Screen_id[infunc_drugName_id]
      infunc_drugName_screenID_trimmed=substring(infunc_drugName_screenID, 0, 3)
      # if single MTS vs single HTS
      if(sum(infunc_drugName_screenID_trimmed=='HTS')==1 &
         sum(infunc_drugName_screenID_trimmed=='MTS')==1){
        infunc_response=infunc_response[infunc_drugName_screenID_trimmed=='MTS',]
        # if HTS vs multiple MTS
      } else if(sum(infunc_drugName_screenID_trimmed=='HTS')==1 &
                sum(infunc_drugName_screenID_trimmed=='MTS')>1){
        infunc_response=infunc_response[infunc_drugName_screenID_trimmed=='MTS',]
        infunc_response[count_row_NAs(infunc_response)==min(count_row_NAs(infunc_response)),]
        # If above the above is still a tie btw MTS screens; Take a mean
        if(is.matrix(infunc_response)){ infunc_response=rowMeans(infunc_response, na.rm = T) }
        # if multiple MTS are competing
      } else if(sum(infunc_drugName_screenID_trimmed=='HTS')==0 &
                sum(infunc_drugName_screenID_trimmed=='MTS')>1){
        infunc_response=infunc_response[
          count_row_NAs(infunc_response)==min(count_row_NAs(infunc_response)),]
        # If above the above is still a tie btw MTS screens; Take a mean
      }
      if(is.matrix(infunc_response)){ infunc_response=colMeans(infunc_response, na.rm = T) }
    }
  } else if(resp_measure_mode=='IC50') {
    # infun_response_matrix=alShabi_drugs_ic50
    # use GDSC to train instead of PRISM
    infun_response_matrix=gdsc1_DOI_IC50
    infunc_response=infun_response_matrix[match(infunc_drugName,
                                                rownames(infun_response_matrix)),]
  } else if(resp_measure_mode=='VFC') {
    # There are curr (12th Feb), there are three drug concentrations 2 use
    # fda_drugs_VFC_atCaverage_5days
    # fda_drugs_VFC_atCMin_5Days
    # fda_drugs_VFC_atCMax
    
    infun_response_matrix=fda_drugs_VFC_atCaverage_5days
    infunc_response=infun_response_matrix[match(infunc_drugName,
                                                rownames(infun_response_matrix)),]
  }
  # Contrasting to the previous, we want to remove 
  # the cell lines which are present in scRNA-seq
  if(infunc_cancerType=='PanCan'){
    CellLines_with_drugResponse=names(infunc_response)
    CellLines_with_bulkExp=colnames(onTarget$expression_rnorm)
    CellLines_with_singleCellExp=colnames(onTarget$scRNA_complete)
    # Exclude 'CellLinee2exclude' cell lines with scRNA-seq
    CellLinee2exclude=onTarget$annotation_20Q4$DepMap_ID[
      grep(exclude_cancer,onTarget$annotation_20Q4$lineage)]
    if(exclude_cancer=='PanCan'){
      CellLinee2exclude=intersect(CellLines_with_singleCellExp, onTarget$annotation_20Q4$DepMap_ID)
    }
    cellLines2remove=intersect(CellLinee2exclude,CellLines_with_singleCellExp)
    
    common_cellLines=setdiff(intersect(CellLines_with_drugResponse, CellLines_with_bulkExp),
                             cellLines2remove)
    print(length(common_cellLines))
  } else{
    infunc_cancerType='lung'
    CellLines_lung=onTarget$annotation_20Q4$DepMap_ID[
      grep(infunc_cancerType,onTarget$annotation_20Q4$lineage)]
    CellLines_with_drugResponse=names(infunc_response)
    CellLines_with_bulkExp=colnames(onTarget$expression_rnorm)
    CellLines_with_singleCellExp=colnames(onTarget$scRNA_complete)
    length(CellLines_with_singleCellExp)
    
    common_cellLines=setdiff(Reduce(intersect, list(CellLines_with_drugResponse,
                                            CellLines_with_bulkExp,
                                            CellLines_lung)),
                             colnames(onTarget$scRNA_complete))
    print(length(common_cellLines))
  }
  
  infunc_response_matchedSubset=unlist(infunc_response[common_cellLines])
  # Pearson correlation is computed
  # Spearman cor can be compute using providing a ranked vector
  cor_profile=t(apply(onTarget$expression_rnorm[infunc_GOI,common_cellLines], 1, function(x) 
    unlist(cor.test_trimmed_v0(x, infunc_response_matchedSubset, method='p')) ))
  cor_profile=cor_profile[order(abs(cor_profile[,2]), decreasing = T),]
  return(cor_profile)
}

# Function to run a parallel version of "feature_ranking_bulk" function.
run_parallel_feature_ranking_bulk<-function(id_cancerType=cancerType,
                                            infunc_DrugsToUse=DOI,
                                            resp_measure_mode='AUC',
                                            infunc_exclude_cancer='lung',
                                            infunc_GOI=genesUsed_toBuild){
  infun_AUC_mode=resp_measure_mode
  featuresRank_fromBulk=mclapply(1:length(infunc_DrugsToUse), function(x)
    err_handle(feature_ranking_bulk(infunc_drugName=infunc_DrugsToUse[x],
                                    infunc_cancerType=id_cancerType,
                                    resp_measure_mode = infun_AUC_mode,
                                    exclude_cancer=infunc_exclude_cancer,
                                    infunc_GOI=genesUsed_toBuild)), mc.cores = 4
  )
  featuresRank_fromBulk
}
```
<!-- Step 2: Using features selected from bulk-expression, we build a model in bulk & then tune the parameters based on performance in sc-rna-seq-->
```{r}
build_on_BULK_v2<-function(infunc_drugName=DOI[1],
                        infunc_cancerType='PanCan',
                        infunc_features=featuersNames_bulk_PanCan[[1]],
                        single_best=featuersNames_bulk_PanCan[[1]][1],
                        k_features=100,
                        mode='AUC',
                        infunc_alpha=1,
                        model_type='glmnet',
                        exclude_cancer='PanCan',
                        alpha_gradient=0.05,
                        lambda_gradient=20,
                        cv_method='cv'){
  # which measure mode
  infunc_response=get_response_matrix(infunc_mode=mode,
                                      infunc_drugName = infunc_drugName,
                                      exclude_cancer=exclude_cancer)
  # which cell lines 2 use during training
  common_cellLines=get_cellLine_list(infunc_cancerType=infunc_cancerType,
                                     infunc_drugName = infunc_drugName,
                                     exclude_cancer=exclude_cancer,
                                     infunc_response = infunc_response,
                                     force_add_cellLines = T)[[1]]
  # which cell lines 2 exclude during training
  cellLines2remove=get_cellLine_list(infunc_cancerType=infunc_cancerType,
                                     infunc_drugName = infunc_drugName,
                                     exclude_cancer=exclude_cancer,
                                     infunc_response = infunc_response)[[2]]
  # whether consistent rnorm
  if(nrow(onTarget$expression_rnorm)!=nrow(onTarget$scRNA_subset_rnorm)){return('Mismatch in initial features\n
                                               rNorm sc vs bulk')}
  ############
  # Train Data:: on bulk
  ##########
  Train_infunc_features_id=head(na.omit(match(infunc_features,
                                             rownames(onTarget$expression_rnorm))), k_features)
  Train_features=t(onTarget$expression_rnorm[Train_infunc_features_id, common_cellLines])
  Train_Target_Label=infunc_response[common_cellLines]
  Train_features=Train_features[which(!is.na(Train_Target_Label)),]
  Train_Target_Label=unlist(Train_Target_Label[which(!is.na(Train_Target_Label))])
  Train_features=Train_features[which(Train_Target_Label<10^100),]
  Train_Target_Label=unlist(Train_Target_Label[which(Train_Target_Label<10^100)])
  ############
  # Test Data 1:: on pseudo-Bulk
  ##########
  cell_ID_for_test_data1=Reduce(intersect,list(cellLines2remove,
                                               colnames(onTarget$scRNA_subset_rnorm),
                                               names(infunc_response) ))
  Test_infunc_features_id=head(na.omit(match(infunc_features,
                                             rownames(onTarget$scRNA_subset_rnorm))), k_features)
  Test_features=t(onTarget$scRNA_subset_rnorm[Test_infunc_features_id, cell_ID_for_test_data1])
  Test_Target_Label=unlist(infunc_response)[cell_ID_for_test_data1]
  Test_features=Test_features[which(!is.na(Test_Target_Label)),]
  Test_Target_Label=unlist(Test_Target_Label[which(!is.na(Test_Target_Label))])
  ############
  # Test Data 2:: on bulk
  ##########
  cell_ID_for_test_data2=Reduce(intersect,list(cellLines2remove,
                                               colnames(onTarget$expression_rnorm),
                                               names(infunc_response) ))
  Test2_infunc_features_id=head(na.omit(match(infunc_features,
                                            rownames(onTarget$expression_rnorm))),
                               k_features)
  Test2_features=t(onTarget$expression_rnorm[Test2_infunc_features_id,cell_ID_for_test_data2])
  Test2_Target_Label=unlist(infunc_response)[cell_ID_for_test_data2]
  Test2_features=Test2_features[which(!is.na(Test2_Target_Label)),]
  Test2_Target_Label=unlist(Test2_Target_Label[which(!is.na(Test2_Target_Label))])
  ############
  # Test Data 3:: on sc-RNA-seq (All cancer)
  ##########
  cells_ID_for_test_data3=onTarget$metadata_CPM_scRNA$NAME[
    onTarget$metadata_CPM_scRNA$DepMap_ID %in% Reduce(intersect,list(cellLines2remove,
                                               names(infunc_response)))]
  cellLine_ID_for_test_data3=unique(
    onTarget$metadata_CPM_scRNA$DepMap_ID[onTarget$metadata_CPM_scRNA$DepMap_ID %in% Reduce(intersect,list(cellLines2remove,names(infunc_response)))])
  Test3_infunc_features_id=head(na.omit(match(infunc_features,
                                            rownames(onTarget$CPM_scRNA_CCLE_rnorm))),
                               k_features)
  Test3_features=t(onTarget$CPM_scRNA_CCLE_rnorm[Test3_infunc_features_id,
                                                 cells_ID_for_test_data3])
  Test3_Target_Label=unlist(infunc_response)[as.character(cellLine_ID_for_test_data3)]
  ##########
  # ML Model
  ##########
  # Non-Linear random Forest based ML model
  # Core Implementation of ATLANTIS (Method Summary: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5667678/)
  # Github: https://github.com/cancerdatasci/atlantis/tree/master/atlantis
  # Note 1: All haematopeotic stem cells are removed from cell lines set used
  # Note 2: Instead of multiple RF runs; ranking features first and providing the top ones
  # for a lower time complexity
  # Note 3: Conditional Inference trees are *not* used,
  # as all features (exp) are continous (no Mut data or categorical)
  # Note 4: Out-of-bag R2 is used as the goodness-of-fit metric
  # Note 5: Non-Goldilock hyperparameters are not optimized
  # Note 6: Nodesize is default and hasn't been tuned (space for improvement)
  set.seed(1)
  tc <- trainControl(method=cv_method,number=3)
  # tc <- trainControl(method='oob')
  # mtry is number of features provided to each split; Tuned here from 10 instances
  if(model_type=='rf'){
  tunegrid <- expand.grid(.mtry=ncol(Train_features)/seq(10, 1, -3))
  cv.out <- train(Train_Target_Label~., # Train_Target_Label is a function of rest of the variables
                  data = data.frame(Train_Target_Label, Train_features), # Use the train data frame as the training data
                  method = 'rf', # Use the 'random forest' algorithm
                  trControl = tc,
                  tunegrid=tunegrid,
                  ntree=500 # Finding Plateue w.r.t this paramter could help sig improve time complexity
                  )
  # GLmnet implementation
  } else if(model_type=='glmnet'){
    tunegrid <- expand.grid(alpha = seq(0,1, alpha_gradient),
                          lambda = seq(0.0001, 1, length = lambda_gradient))
    cv.out <- train(Train_Target_Label~., # Train_Target_Label is a function of rest of the variables
                  data = data.frame(Train_Target_Label, Train_features), # Use the train data frame as the training data
                  method = 'glmnet',
                  trControl = tc,
                  metric="RMSE",
                  tuneGrid=tunegrid,
                  savePredictions="final")
  }
  ############
  # Performance
  ##########
  # Best performance while tunig parameters
  model_performance_cv=max(sqrt(cv.out$results$Rsquared), na.rm = T)
  # Test on bulk (test 1)
  Test_pred_bulk=predict(cv.out, Test2_features)
  performance_in_bulk=unlist(cor.test(Test_pred_bulk,
                                          Test2_Target_Label, method='p')[c(3, 4)])
  # Test on pseudo-Bulk (test 2)
  Test_pred_Msc=predict(cv.out, newdata = Test_features)
  performance_in_pseudo_bulk=unlist(cor.test(Test_pred_Msc,
                                         Test_Target_Label, method='p')[c(3, 4)])
  # Test on scRNA-seq:: (test 3)
  Test_pred_sc=predict(cv.out, newdata = Test3_features)
  cells_mapping2_cellLine=onTarget$metadata_CPM_scRNA$DepMap_ID[
    onTarget$metadata_CPM_scRNA$NAME %in% names(Test_pred_sc)]
  Test_pred_sc_mean_byCellLine=aggregate(Test_pred_sc ~cells_mapping2_cellLine, 
                                         data.frame(Test_pred_sc, cells_mapping2_cellLine),
                                         function(x) mean(x, na.rm=T))
  performance_in_scRNA=unlist(cor.test(Test_pred_sc_mean_byCellLine$Test_pred_sc,
                                         Test3_Target_Label[
                                           as.character(
                                             Test_pred_sc_mean_byCellLine$cells_mapping2_cellLine)], method='p')[c(3, 4)])
  # To reutrn the predictions
  predVSgroundTruth=list(
    # scRNA-seq 
    pred_gt_scRNA=data.frame(Test_pred_sc=Test_pred_sc_mean_byCellLine$Test_pred_sc,
                            Observed=Test3_Target_Label[as.character(Test_pred_sc_mean_byCellLine$cells_mapping2_cellLine)]),
    # bulk-seq
    pred_gt_bulk=data.frame(Test_pred_Msc=Test_pred_bulk,
                            Observed=Test2_Target_Label),
    # for pseudo-bulk-seq test
    pred_gt_mscRNA=data.frame(Test_pred_Msc=Test_pred_Msc,
                              Observed=Test_Target_Label)
    )
  
  ############
  # object 2 return
  ##########
  toreturn=list(model=cv.out,
                single_best=single_best,
                model_performance_during_cv=model_performance_cv,
                performance_in_bulk=performance_in_bulk,
                performance_in_pseudo_bulk=performance_in_pseudo_bulk,
                performance_in_scRNA=performance_in_scRNA,
                predVSgroundTruth=predVSgroundTruth)
  toreturn
}

# Running the above function in parallel
build_biomarker_step2_bulk<-function(features_fromBulk=featuers_bulk_all_cancer_type,
                                     id_cancerType=cancerType,
                                     infunc_drugsList=DOI,
                                     resp_measure_mode='AUC',
                                     infunc_k_features,
                                     infunc_alpha=1,
                                     infunc_model_type='glmnet',
                                     infunc_exclude_cancer){
  infunc_drugsList=as.character(infunc_drugsList)
  model_for=lapply(1:length(infunc_drugsList),
                     function(x) 
                       err_handle(build_on_BULK_v2(infunc_drugName=infunc_drugsList[x],
                                                infunc_cancerType=id_cancerType,
                                                infunc_features=rownames(features_fromBulk[[x]]),
                                                single_best=features_fromBulk[[x]][1,],
                                                k_features=infunc_k_features,
                                                mode=resp_measure_mode,
                                                model_type=infunc_model_type,
                                                exclude_cancer=infunc_exclude_cancer)))
  model_for
}

# Outputs various performance measure using bulk, pseudo-bulk, single-cell during training and testing
compare_performance <- function(model_for){
infunc_compare_performance=list(model_perf_CV=do.call(rbind, lapply(model_for,
                                                               function(x) err_handle(x[[3]]))),
                           model_perf_test_bulk=do.call(rbind, lapply(model_for,
                                                                 function(x) err_handle(x[[4]]))),
                           model_perf_test_mean_scRNA=do.call(rbind, lapply(model_for,
                                                                      function(x) err_handle(x[[5]]))),
                           model_perf_test_actual_scRNA=do.call(rbind, lapply(model_for,
                                                                      function(x) err_handle(x[[6]]))),
                           model_perf_test_mean_allCancer=do.call(rbind, lapply(model_for,
                                                                      function(x) err_handle(x[[7]])))
                           )
  # print('Number of Drugs passing a lenient Threshold of model performance')
  print(sum(infunc_compare_performance[[1]][,1]>0.3, na.rm = T))  
  print(sapply(infunc_compare_performance, function(x)
    err_handle(sum(x[,2]>0.3, na.rm = T) )) )
  print(sapply(infunc_compare_performance, function(x)
    err_handle(mean(x[,2], na.rm = T) )) )
  print(DOI[which(infunc_compare_performance$model_perf_test_actual_scRNA[,2]>0.3)])
  
  df2ret=lapply(infunc_compare_performance, function(x)
    err_handle(which(x[,1]>0.3) ))[[4]]
  sapply(infunc_compare_performance, function(x)
    err_handle(mean(x[,2], na.rm = T) ))
}
```
<!-- get Cell lines list and response matrix of a particular cancer type considering exclusions-->
```{r}
get_response_matrix <- function(infunc_mode=mode,
                                infunc_drugName, exclude_cancer){
  if(infunc_mode=='AUC'){
    # Mode 1: All HTS
    # Mode 2: Priortize MTS
    infun_response_matrix=onTarget$secondary_prism
    infunc_drugName_id=which(stripall2match(onTarget$secondary_screen_drugAnnotation$CommonName)==infunc_drugName)
    infunc_response=onTarget$secondary_prism[infunc_drugName_id,]
    # is.matrix(infunc_response) - denotes if there exists a matrix with multiple screen
    # If there are multiple measures of response;
    if(is.matrix(infunc_response)){
      # Choose screen of interest
      # Rule: Priority, MTS over HTS; MTS with least number of NAs
      infunc_drugName_screenID=onTarget$secondary_screen_drugAnnotation$Screen_id[infunc_drugName_id]
      infunc_drugName_screenID_trimmed=substring(infunc_drugName_screenID, 0, 3)
      # if single MTS vs single HTS
      if(sum(infunc_drugName_screenID_trimmed=='HTS')==1 &
         sum(infunc_drugName_screenID_trimmed=='MTS')==1){
        infunc_response=infunc_response[infunc_drugName_screenID_trimmed=='MTS',]
        # if HTS vs multiple MTS
      } else if(sum(infunc_drugName_screenID_trimmed=='HTS')==1 &
                sum(infunc_drugName_screenID_trimmed=='MTS')>1){
        infunc_response=infunc_response[infunc_drugName_screenID_trimmed=='MTS',]
        infunc_response[count_row_NAs(infunc_response)==min(count_row_NAs(infunc_response)),]
        # If above the above is still a tie btw MTS screens; Take a mean
        if(is.matrix(infunc_response)){ infunc_response=rowMeans(infunc_response, na.rm = T) }
        # if multiple MTS are competing
      } else if(sum(infunc_drugName_screenID_trimmed=='HTS')==0 &
                sum(infunc_drugName_screenID_trimmed=='MTS')>1){
        infunc_response=infunc_response[
          count_row_NAs(infunc_response)==min(count_row_NAs(infunc_response)),]
        # If above the above is still a tie btw MTS screens; Take a mean
      }
      if(is.matrix(infunc_response)){ infunc_response=colMeans(infunc_response, na.rm = T) }
    }
  } else if(infunc_mode=='IC50'){
    # infun_response_matrix=alShabi_drugs_ic50
    # use GDSC to train instead of PRISM
    infun_response_matrix=gdsc1_DOI_IC50
    infunc_response=infun_response_matrix[match(infunc_drugName,
                                                rownames(infun_response_matrix)),]
  } else{
    # There are curr (12th Feb), there are three drug concentrations 2 use
    # fda_drugs_VFC_atCaverage_5days
    # fda_drugs_VFC_atCMin_5Days
    # fda_drugs_VFC_atCMax
    
    infun_response_matrix=fda_drugs_VFC_atCaverage_5days
    infunc_response=infun_response_matrix[match(infunc_drugName, rownames(infun_response_matrix)),]
  }
  infunc_response
}

get_cellLine_list <- function(infunc_cancerType=infunc_cancerType, infunc_drugName, exclude_cancer,
                              infunc_response=infunc_response,
                              force_add_cellLines=T,
                              force_add_cellLines_list=NA){
  if(infunc_cancerType=='PanCan'){
    CellLines_with_drugResponse=names(infunc_response)
    CellLines_with_bulkExp=colnames(onTarget$expression_rnorm)
    CellLines_with_singleCellExp=colnames(onTarget$scRNA_complete)
    
    # Exclude 'CellLinee2exclude' cell lines with scRNA-seq
    CellLinee2exclude=onTarget$annotation_20Q4$DepMap_ID[
      grep(exclude_cancer,onTarget$annotation_20Q4$lineage)]
    if(exclude_cancer=='PanCan'){
      CellLinee2exclude=intersect(CellLines_with_singleCellExp, onTarget$annotation_20Q4$DepMap_ID)
    }
    cellLines2remove=intersect(CellLinee2exclude,CellLines_with_singleCellExp)
    
    cellLines2remove_allCancer=intersect(onTarget$annotation_20Q4$DepMap_ID,CellLines_with_singleCellExp)
    # common_cellLines=setdiff(intersect(CellLines_with_drugResponse, CellLines_with_bulkExp), 
    #         cellLines2remove_allCancer)
    common_cellLines=setdiff(intersect(CellLines_with_drugResponse, CellLines_with_bulkExp), 
                             cellLines2remove)  
    if(force_add_cellLines){
    common_cellLines=c(common_cellLines, intersect(CellLines_with_bulkExp, force_add_cellLines_list))
    }
  } else{
    infunc_cancerType='lung'
    CellLines_lung=onTarget$annotation_20Q4$DepMap_ID[
      grep(infunc_cancerType,onTarget$annotation_20Q4$lineage)]
    CellLines_with_drugResponse=names(infunc_response)
    CellLines_with_bulkExp=colnames(onTarget$expression_rnorm)
    CellLines_with_singleCellExp=colnames(onTarget$scRNA_complete)
    
    # Romve only lung with scRNA-seq
    cellLines2remove=intersect(CellLines_lung,CellLines_with_singleCellExp)
    cellLines2remove_allCancer=intersect(onTarget$annotation_20Q4$DepMap_ID,CellLines_with_singleCellExp)
    common_cellLines=setdiff(Reduce(intersect, list(CellLines_with_drugResponse,
                                            CellLines_with_bulkExp,
                                            CellLines_lung)),
                             cellLines2remove)
  }
  list(common_cellLines, cellLines2remove)
}
```
<!-- # Build funcs to predict viability -->
```{r}
# Given a drug response model and an expression matrix, this function returns a viability matrix
viability_from_model<-function(infunc_DruName,
                               infun_model,
                               infunc_dataset){
  # FOI stands for features of interest
  infunc_dataset_FOI=infunc_dataset
  infunc_dataset_FOI_t=data.frame(t(infunc_dataset_FOI))
  
  Viability_score=predict(infun_model,
                          infunc_dataset_FOI_t)
  Viability_score
}

killing_in_each_dataset<-function(infunc_scRNAseq_dataset_rnorm,
                                  infunc_GOI=genesUsed_toBuild,
                                  infun_model_list=Pred_models_list){
  x=1
  viab_raw=lapply(1:length(infun_model_list), function(x) 
    viability_from_model(infunc_DruName=names(infun_model_list[x]),
                         infun_model=infun_model_list[[x]]$model,
                         infunc_dataset=infunc_scRNAseq_dataset_rnorm))
  names(viab_raw)=names(infun_model_list)
  do.call(cbind, viab_raw)
}
each_patient_killing <- function(x=1,
                                 mode= 'weighted_average',
                                 clone_killing_matrix=mono_killing_df){
  total_cells=sum(Clone_Counts_per_patients[x,clone_killing_matrix[
    clone_killing_matrix$patient==Clone_Counts_per_patients$patients[x],]$clone_id])
  clone_weights=unlist(Clone_Counts_per_patients[x,clone_killing_matrix[
    clone_killing_matrix$patient==Clone_Counts_per_patients$patients[x],]$clone_id]/total_cells)
  if(mode=='weighted_average'){
    weighted_killing=sum(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],]$comb_killing*clone_weights)
  } else if(mode=='min'){
    weighted_killing=min(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],]$comb_killing)
  } else if(mode=='max'){
    weighted_killing=max(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],]$comb_killing)
  } else if(mode=='weighted_max'){
    weighted_killing=max(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],]$comb_killing*clone_weights)
  }
  weighted_killing
}
each_patient_killingv2 <- function(x=1,
                                   mode= 'weighted_average',
                                   clone_killing_matrix=mono_killing_df){
  total_cells=sum(Clone_Counts_per_patients[x,clone_killing_matrix[
    clone_killing_matrix$patient==Clone_Counts_per_patients$patients[x],]$clone_id])
  clone_weights=unlist(Clone_Counts_per_patients[x,clone_killing_matrix[
    clone_killing_matrix$patient==Clone_Counts_per_patients$patients[x],]$clone_id]/total_cells)
  if(mode=='weighted_average'){
    killing=apply(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],-(1:2)], 2, function(x) 
                                       sum(x*clone_weights) )
  } else if(mode=='min'){
    killing=apply(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],-(1:2)], 2, function(x) 
                                       min(x) )
  } else if(mode=='max'){
    killing=apply(clone_killing_matrix[clone_killing_matrix$patient==
                                     Clone_Counts_per_patients$patients[x],-(1:2)], 2, function(x) 
                                       max(x) )
  }
  killing
}
```
<!-- Read models -->
```{r}
# THis function is used to load only the performance of a model
# Could be used in case one is accessing a large number of models performance and have limited memory.
get_performance <- function(infunc_FOI=fileNames_ofInterest[1]){
  readRDS(paste('Data/cancer_approved_drugs/', infunc_FOI, sep=''))[3:7]
}
```
<!-- Compute meta-information for clones -->
```{r}
# <!-- compute clone abundance proportion (weight) in each patient -->
each_patient_clone_weights <- function(x=1, mode= 'weighted_average'){
  total_cells=sum(Clone_Counts_per_patients[x,comb_killing_df[
    comb_killing_df$patient==Clone_Counts_per_patients$patients[x],]$clone_id])
  clone_weights=unlist(Clone_Counts_per_patients[x,comb_killing_df[
    comb_killing_df$patient==Clone_Counts_per_patients$patients[x],]$clone_id]/total_cells)
  initialize_clone_weights=c(c1=0, c2=0, c3=0)
  initialize_clone_weights[match(names(clone_weights), names(initialize_clone_weights))]=clone_weights
  initialize_clone_weights
}
# <!-- each_patient_pseudo_bulk -->
each_patient_pseudo_bulk <- function(x=1){
  total_cells=sum(Clone_Counts_per_patients[x,comb_killing_df[
    comb_killing_df$patient==Clone_Counts_per_patients$patients[x],]$clone_id])
  clone_weights=unlist(Clone_Counts_per_patients[x,comb_killing_df[
    comb_killing_df$patient==Clone_Counts_per_patients$patients[x],]$clone_id]/total_cells)
  clone_expression=data.frame(clone_Level_z_expression_df[,grep(Clone_Counts_per_patients$patients[x], colnames(clone_Level_z_expression_df))])
  if(ncol(clone_expression)>1){
  pseudo_bulk=rowMeans(clone_Level_z_expression_df[,grep(Clone_Counts_per_patients$patients[x], colnames(clone_Level_z_expression_df))]*clone_weights)
  } else{
    pseudo_bulk=clone_Level_z_expression_df[,grep(Clone_Counts_per_patients$patients[x], colnames(clone_Level_z_expression_df))]
  }
  pseudo_bulk
}
```
<!-- Calculate viability from a model at sc-level -->
```{r}
viability_from_model<-function(infunc_DruName=names(approved_drugs_model_for_MM_patients)[1],
                               infun_model=approved_drugs_model_for_MM_patients[[1]]$model,
                               infunc_dataset=
                                 onTarget$CPM_scRNA_CCLE_rnorm[
      match(approved_drugs_model_for_MM_patients[[1]]$model$coefnames, 
            rownames(onTarget$CPM_scRNA_CCLE_rnorm)),
      needed_cells_id]
      ){
  # FOI stands for features of interest
  
  infunc_dataset[match(infun_model$coefnames,rownames(infunc_dataset)),]
  infunc_dataset_FOI=data.frame(infunc_dataset[match(infun_model$coefnames,
                                                     rownames(infunc_dataset)),],
                                row.names = rownames(infunc_dataset)[match(infun_model$coefnames,
                                                     rownames(infunc_dataset))])
  infunc_dataset_FOI=data.frame(t(infunc_dataset_FOI))
  Viability_score=predict(infun_model, infunc_dataset_FOI)
  Viability_score
}
```
<!-- seurat_clustering given a expression matrix -->
```{r}
seurat_clustering <- function(input_matrix){
  lung_so <- CreateSeuratObject(counts = input_matrix,
                                project = "lungv0",
                                min.cells = 3,
                                min.features = 200)
  lung_so <- NormalizeData(lung_so, normalization.method = "LogNormalize", scale.factor = 10000)
  lung_so <- FindVariableFeatures(lung_so, selection.method = "vst", nfeatures = 2000)
  lung_so <- ScaleData(lung_so)
  lung_so <- RunPCA(lung_so, features = VariableFeatures(object = lung_so))
  lung_so <- FindNeighbors(lung_so, dims = 1:10)
  lung_so <- FindClusters(lung_so, resolution = 0.8)
  cluster_id=Idents(lung_so)
  return(cluster_id)
}
```
<!-- Genes used to build models -->
```{r}
genesUsed_toBuild=intersect(rownames(onTarget$expression_20Q4),
                            rownames(onTarget$scRNA_complete))
# saveRDS(genesUsed_toBuild, 'Data/genesUsed_toBuild.RDS')
```

```{r}
# genes_across_scRNA_datasets_ofInterest=Reduce(intersect, list(rownames(onTarget$scRNA_complete),
#                                                               rownames(onTarget$expression_20Q4)))
# saveRDS(genes_across_scRNA_datasets_ofInterest, 'Data/genes_across_scRNA_datasets_ofInterest.RDS')
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.